---
title: "Introduction to rPinecone"
author: "Alex Wailan"
date: "`r Sys.Date()`"
bibliography: bibliography.bib
output: 
  html_document:
    fig_width: 12
    fig_height: 8
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8,
                      echo=TRUE, warning=FALSE, message=FALSE,
                      tidy=TRUE)
```

rPinecone, is an R package designed to define sub-lineages within closely related LV populations. rPinecone uses root-to-tip directional approach to define sub-lineages within a phylogenetic tree according to SNV distance from the ancestral node. 

***


## Libraries

```{r}
library(phytools)
library(rPinecone)
```

The package defines sub-lineages within a bacterial clonal expansion via a phylogenetic tree.

## Inputs

* A rooted tree in newick format with branch lengths representing SNP distance

    A SNP scaled tree can be produced using the algorithm of Pupko et al. available in python at [pyjar](https://github.com/simonrharris/pyjar#usage)

* SNP threshold - used for sub-grouping
* Relatability threshold - the number of internal vertices each sub-lineage must have between each other to form a Major Sub-group

## Operation

After loading a newick formatted tree rPinecone can be called as

```{r}
tree.file.name <- system.file("extdata", "pyjar.staph_ST2371_45_re_itol_7079_1_6_root.joint.tre", package = "rPinecone")
tree <- phytools::read.newick(tree.file.name)
results <- pinecone(tree, 2, 3, quiet = TRUE)
```

## Exporting results

The resulting output can the be parsed for plotting with iTOL

To exports a data file for replacing Tip Labels with Sub-Group number (LABELS) run

```{r}
itol_labels_template(results)
```

To export a data file for displaying the Sub-Groups (DATASET_COLOURSTRIP) run
```{r}
itol_sublineage_output(results)
```

To export a data file for displaying the Major Sub-Groups (DATASET_COLOURSTRIP) run
```{r}
itol_major_SL_output(results)
```

The phylogentic tree with dichotomies resolved into multichotomies can be saved by running
```{r}
write.tree(results$tree, file = "rpinecone.tree")
```

## References

---
nocite: '@*'
...

##Appendix

```{r}
sessionInfo()
```
